Investigating magnetic dynamo evolution with \TESS\ field dwarfs

### 2017-10-06 ms.tex
blah

\ssfigure{Davenport.png}{0.5}{%
From Davenport (2017).
    Rotation period as a function of effective
temperature for the full McQuillan et al., (2014) \Kepler\ sample in black, and
the subset of these stars that also feature in the TGAS Gaia DR1 catalogue in
blue. Contaminating giants have been removed from the blue sample and the
rotation period bimodality is revealed to exist across all temperatures shown.
    The red line is a 600 Myr rotational isochrone (also known as a gyrochrone).
\label{fig:davenport}
}

Although the classical spin-down law of \citet{skumanich1972} (Period
$\propto$ Age$^{1/2}$) holds for all open clusters with measured periods, it
does not appear to describe old field stars \citep{angus2015, van-saders2016}.
\citet{van-saders2016} found that including a transition to a weakened magnetic
braking regime in the gyrochronology models, at a critical Rossby number,
provides an improved fit to the data.
Further calibration is needed however; a lack of rotation periods and reliable
ages for old and low-mass stars leaves the rotational evolution of stars older
than the Sun relatively unexplored.
\TESS\ will provide thousands of new {\it long } rotation periods for stars
that fall in the CVZ.
Many of these will have spectra from the \TESS --\HERMES\ survey, from which
it will be possible to calculate isochronal ages, enabling further calibration
of the gyrochronology relations.
{\bf We propose to calibrate gyrochronology at old ages for FGKM stars using
the isochronal ages of \TESS--\hermes\ stars.}
We will also use the rotation periods of \TESS\ field dwarfs over the whole sky
to test the gyrochronology relations using galactic kinematics.
Many \tess\ FFI targets will have proper motions, parallaxes, positions and
radial velocities published in the second \Gaia\ data release.
These parameters provide the information necessary to calculate galactocentric
positions and action-angle coordinates of the stars, both of which are age
indicators.
{\bf We propose to calibrate the gyrochronology relations at old ages for FGKM
stars using vertical actions of \Gaia\ DR2 targets.}
Since one of the few ways to accurately age-date fully convective, late M
dwarfs is via kinematics, these new relations will help to infer ages for all
stars with M$<$ 0.35M$_\odot$, to which gyrochronology cannot be applied.

@@

Based on \citet{sullivan2015}, we expect around 10 million \TESS\ FFI
targets brighter than 15th magnitude in the southern ecliptic hemisphere.
We intend to prioritize certain groups of stars for period analysis,
calculating a simple (and computationally fast) Lomb-Scargle periodogram for
the low priority objects and running a full probabilistic, Gaussian process
period analysis on the high priority targets.
High priority targets will include: comoving pairs, stars in the CVZ
(especially those with spectra), stars in TGAS, stars at large distances, plus
potential asteroseismic dwarfs and exoplanet hosts.
The number of high priority targets with probabilistic rotation periods will
be limited by computation power, but we conservatively estimate that we can
produce full rotation period posterior PDFs for $\sim$ 30,000 stars.
Based on the expected photometric precision of \tess\ \citep{sullivan2015}, we
expect to be able to measure rotation periods for FFI stars down to 15th
magnitude.
We note however that not all stars will have measurable rotation periods:
even if they are bright enough not all stars show rotational variability in
their light curves either because they are pole-on or inactive.
We will vet incorrect rotation period measurements using a combination of a
Lomb-Scargle periodogram and an autocorrelation function by
training on a set of light curves with a range of variability amplitudes and
known periods.
Based on the fraction of stars with measurable rotation periods in
\citet{mcquillan2014}, and the more restrictive baseline of \tess\ relative to
\kepler, we predict that the final catalog will contain around one million
rotation periods.

>>>>

\ssfigure{Davenport.png}{0.5}{%
From Davenport (2017).
    Rotation period as a function of effective temperature for the full
McQuillan et al., (2014) \Kepler\ sample in black, and the subset of these
stars that also feature in the Tycho-Gaia Astrometric Solution (TGAS) Gaia DR1
catalog in blue. Contaminating giants have been removed from the blue sample
and the
rotation period bimodality is revealed to exist across all temperatures shown.
    The red line is a 600 Myr rotational isochrone (also known as a gyrochrone).
\label{fig:davenport}
}

Although the classical spin-down law of \citet{skumanich1972} (Period
$\propto$ Age$^{1/2}$) holds for all open clusters with measured periods, it
does not appear to describe old field stars \citep{angus2015, van-saders2016}.
\citet{van-saders2016} found that including a transition to a weakened magnetic
braking regime in the gyrochronology models, at a critical Rossby number,
provides an improved fit to the data.
Further calibration is needed however; a lack of rotation periods and reliable
ages for old and low-mass stars leaves the rotational evolution of stars older
than the Sun relatively unexplored.
\TESS\ will provide thousands of new {\it long } rotation periods for stars
that fall in the CVZs.
Many of these will have spectra from the \TESS --\HERMES\ survey, from which
it will be possible to calculate isochronal ages, enabling further calibration
of the gyrochronology relations.
{\bf We propose to calibrate gyrochronology at old ages for FGKM stars using
the isochronal ages of \TESS--\hermes\ stars.}
We will also use the rotation periods of \TESS\ field dwarfs over the whole sky
to test the gyrochronology relations using galactic kinematics.
Many \tess\ FFI targets will have proper motions, parallaxes, positions and
radial velocities published in the second \Gaia\ data release.
These parameters provide the information necessary to calculate galactocentric
positions and action-angle coordinates of the stars, both of which are age
indicators.
{\bf We propose to calibrate the gyrochronology relations at old ages for FGKM
stars using vertical actions of \Gaia\ DR2 targets.}
Since one of the few ways to accurately age-date fully convective, late M
dwarfs is via kinematics, these new relations will help to infer ages for all
stars with M$<$ 0.35M$_\odot$, to which gyrochronology cannot be applied.

@@

Based on \citet{sullivan2015}, we expect around 10 million \TESS\ FFI
targets brighter than 15th magnitude in the southern ecliptic hemisphere.
We intend to prioritize certain groups of stars for period analysis,
calculating a simple (and computationally fast) Lomb-Scargle periodogram for
the low priority objects and running a full probabilistic, Gaussian process
period analysis on the high priority targets.
High priority targets will include: comoving pairs, stars in the CVZ
(especially those with spectra), stars in TGAS, stars at large distances, plus
potential asteroseismic dwarfs and exoplanet hosts.
The number of high priority targets with probabilistic rotation periods will
be limited by computation power, but we conservatively estimate that we can
produce full rotation period posterior PDFs for $\sim$ 30,000 stars.
Based on the expected photometric precision of \tess\ \citep{sullivan2015}, we
expect to be able to measure rotation periods for FFI stars down to 15th
magnitude.
We note however that not all stars will have measurable rotation periods:
even if they are bright enough not all stars show rotational variability in
their light curves either because they are pole-on or inactive.
We will vet incorrect rotation period measurements by training our algorithms
on a set of light curves with a range of variability amplitudes and known
periods.
Based on the fraction of stars with measurable rotation periods in
\citet{mcquillan2014}, and the more restrictive baseline of \tess\ relative to
\kepler, we predict that the final catalog will contain around one million
rotation periods.
